---
title: "The 'gmpoly' package"
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/stla/gmpoly/workflows/R-CMD-check/badge.svg)](https://github.com/stla/gmpoly/actions)
<!-- badges: end -->

*R package for multivariate polynomials with rational coefficients.*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```

```{r}
library(gmpoly)
```

Define a polynomial with the `gmpoly` function:

```{r}
pol <- gmpoly("4 x^(2, 1, 1) + 1/2 x^(0,1,0)")
pol
```

Some arithmetic on this polynomial:

```{r}
-pol
2 * pol
pol / 2
pol + 5
pol - 5
pol^2
```

Note that you cannot directly use a `bigq` scalar, e.g. you can't do 
`gmp::as.bigq(5, 3) * pol`. To perform such an operation, you have to use 
`gmpolyConstant`:

```{r}
gmpolyConstant(3, "5/3") * pol
```

The `gmpolyConstant` function converts a scalar to a constant polynomial, and 
two polynomials can be added and multiplied:

```{r}
pol1 <- gmpoly("2 x^(1,1) - 5/3 x^(0,1)")
pol2 <- gmpoly("-2 x^(1,1) + 3 x^(2,1)")
pol1 + pol2
pol1 * pol2
```

Use `gmpolyEval` to evaluate a polynomial for some values of the variables:

```{r}
library(gmp, warn.conflicts = FALSE)
pol <- gmpoly("5/2 x^(2,3) + 3 x^(1,1)")
gmpolyEval(pol, x = as.bigq(c(1, 1)))
```

To evaluate the polynomial for several sets of variables, supply a matrix to 
the second argument:

```{r}
X <- rbind(
  t(as.bigq(c(1, 1))),
  t(as.bigq(c(3, 4), c(4, 3)))
)
gmpolyEval(pol, x = X)
```

